Musical Genres and Socio-Demographic Labels as Fuzzy Categories

### 2025-06-21 Old/analysis-bb-old.qmd
reorg

```{r setup, include=FALSE}
   library(backbone)
   library(conflicted)
   library(dplyr)
   library(FactoMineR)
   library(factoextra)
   library(ggplot2)
   library(ggrepel)
   library(corrplot)
   library(ggridges)
   library(haven)
   library(here)
   library(paletteer)
   library(stringr)
   cols <- paletteer_dynamic("cartography::multi.pal", 20)
   conflicts_prefer(dplyr::filter)
   conflicts_prefer(dplyr::select)
   source(here("make.fuzz.plot.R"))
```

```{r import data and select genre stereotype variables}
   #importing data
   dat <- read_dta("C:/Users/Omar Lizardo/Google Drive/MISC DATA SOURCES/SSI-2012/SSI2012.dta")
   #replacing ids with character names
   dat <- bind_cols(tibble(id_char = paste("id", sapply(dat[, 1], as.character), sep = "_")), dat) %>% 
   select(id_char, classicalfemale:hvymtluc, -ends_with("none")) %>%  #selecting relevant columns
   na.omit() #removing missing values
```

```{r Creating individual matrices}
   row.breiger <- function(x) {
      x %*% t(x)
      }
   col.breiger <- function(x) {
      t(x) %*% x
      }
   bb <- function(x) {
      x <- suppressMessages(disparity(x, alpha = 0.25, narrative = FALSE))
      return(x)
      }
   den <- rowSums(dat[,2:301])/300
   bad.den <- which(den > .9 | den < 0.1)
   dat2 <- dat[-c(bad.den), 2:301]
   ind.list <- split(dat2, seq(nrow(dat2))) # splitting each row of data frame into list objects
   ind.mat <- lapply(ind.list, function(x) {matrix(as.numeric(x), 20, 15, byrow = TRUE)}) # creating individual two mode network of genre by labels
      r <- c("Classical", "Opera", "Jazz", "Broadway", "Easy", "Big_Band", "Classic_Rock", "Country", "Bluegrass", "Folk", "Gospel", "Latin", "Rap", "Blues", "Reggae", "Pop", "Contemp_Rock", "Alt_Rock", "Dance", "Metal")
       c <- c("Female", "Male", "White", "Black", "Hispanic", "Asian", "College", "No_College", "Young", "Mid_Age", "Old", "Lower_Class", "Working_Class", "Middle_Class", "Upper_Class")
   ind.mat <- lapply(ind.mat, function(x) {rownames(x) <- r; x}) # naming individual rows
   ind.mat <- lapply(ind.mat, function(x) {colnames(x) <- c; x}) # naming individual columns
   names(ind.mat) <- dat$id_char[-c(bad.den)]
   ind.row <- lapply(ind.mat, row.breiger) # individual genre projection
   ind.col <- lapply(ind.mat, col.breiger) # individual label projection
   row.bb <- lapply(ind.row, bb)
   col.bb <- lapply(ind.col, bb)
   row.stack <- matrix(unlist(row.bb), ncol = 20, byrow = TRUE) # genre stacked matrix
   col.stack <- matrix(unlist(col.bb), ncol = 15, byrow = TRUE) # label stacked matrix
   colnames(row.stack) <- r # naming genre stacked matrix columns
   colnames(col.stack) <- c # naming label stacked matrix columns
   rownames(row.stack) <- rep(r, nrow(dat2)) # naming genre stacked matrix rows
   rownames(col.stack) <- rep(c, nrow(dat2)) # naming label stacked matrix rows
   zero.rows.r <- which(rowSums(row.stack) == 0) # which genre stacked matrix rows sum to zero?
   zero.rows.c <- which(rowSums(col.stack) == 0) # which label stacked matrix rows sum to zero?
   names(ind.row) <- dat$id_char[-c(bad.den)]
   names(ind.col) <- dat$id_char[-c(bad.den)]
   names(row.bb) <- dat$id_char[-c(bad.den)]
   names(col.bb) <- dat$id_char[-c(bad.den)]
```

```{r Stacked CA}
   #genres
   row.stack.d <- data.frame(row.stack)
   row.stack.d$genre <- rep(r, nrow(dat2))
   row.stack.d$id <- rep(dat$id_char[-bad.den], each = 20)
   CA.r <- CA(row.stack.d, graph = FALSE, row.sup = zero.rows.r, quali.sup = c(21, 22))
   #labels
   col.stack.d <- data.frame(col.stack)
   col.stack.d$label <- rep(c, nrow(dat2))
   col.stack.d$id <- rep(dat$id_char[-bad.den], each = 15)
   CA.c <- CA(col.stack.d, graph = FALSE, row.sup = zero.rows.c, quali.sup = c(16, 17))
```

@@

```{r Genre Cloud Plots}
   source(here("make.fuzz.plot.R"))
   h <- 625
   w <- 750
   png(file = here("Plots", "genre-proto-ex1.png"), width=w, height=h)
      make.fuzz.plot(row.gen.dat, col.gen.dat, sup.gen.dat, 
                     seed = 125, ls = 5,
                     1, 2, pa = .35,  t = 3, b = 3, l = 2, r = 3,
                     c("Rap", "Folk", "Classical"))
   dev.off()
   png(file = here("Plots", "genre-proto-ex2.png"), width=w, height=h)
      make.fuzz.plot(row.gen.dat, col.gen.dat, sup.gen.dat,
                     seed = 125, ls = 5,
                     1, 2, pa = .35,  t = 2, b = 2, l = 2, r = 3,
                     c("Alt_Rock", "Gospel", "Opera"))
   dev.off()
   png(file = here("Plots", "genre-proto-ex3.png"), width=w, height=h)
      make.fuzz.plot(row.gen.dat, col.gen.dat, sup.gen.dat,
                     seed = 125, ls = 5,
                     1, 2, pa = .35,  t = 2, b = 2, l = 2, r = 3, bns = 5,
                     c("Metal", "Bluegrass", "Broadway"))
   dev.off()
   png(file = here("Plots", "genre-proto-ex4.png"), width=w, height=h)
      make.fuzz.plot(row.gen.dat, col.gen.dat, sup.gen.dat,
                     seed = 125, ls = 5,
                     1, 2, pa = .35,  t = 2, b = 2, l = 2, r = 3, bns = 4,
                     c("Jazz", "Reggae", "Blues", "Rap", "Classical"))
   dev.off()
   png(file = here("Plots", "genre-proto-ex5.png"), width=w, height=h)
      make.fuzz.plot(row.gen.dat, col.gen.dat, sup.gen.dat,
                     seed = 125, ls = 5,
                     1, 2, pa = .35,  t = 2, b = 2, l = 2, r = 3, bns = 4,
                     c("Country", "Bluegrass", "Classic_Rock", "Folk"))
   dev.off()
   png(file = here("Plots", "genre-proto-ex6.png"), width=w, height=h)
      make.fuzz.plot(row.gen.dat, col.gen.dat, sup.gen.dat,
                     seed = 125, ls = 5,
                     1, 2, pa = .35,  t = 2, b = 2, l = 2, r = 3, bns = 4,
                     c("Alt_Rock", "Contemp_Rock", "Classic_Rock", "Metal"))
   dev.off()
   png(file = here("Plots", "genre-proto-ex7.png"), width=w, height=h)
      make.fuzz.plot(row.gen.dat, col.gen.dat, sup.gen.dat,
                     seed = 125, ls = 5,
                     1, 2, pa = .35,  t = 2, b = 2, l = 2, r = 3, bns = 4,
                     c("Broadway", "Classical", "Opera", "Easy", "Jazz"))
   dev.off()
```

```{r Label Cloud Plots}
   source(here("make.fuzz.plot.R"))
   h <- 625
   w <- 750
   png(file = here("Plots", "label-proto-ex1.png"), width=w, height=h)
      make.fuzz.plot(row.lab.dat, col.lab.dat, sup.lab.dat, bns = 6, ls = 6,
                     1, 2, pa = .35,  t = 2, b = 2, l = 2, r = 2,
                     c("White", "Asian", "Black", "Hispanic"))
   dev.off()
   png(file = here("Plots", "label-proto-ex2.png"), width=w, height=h)
      make.fuzz.plot(row.lab.dat, col.lab.dat, sup.lab.dat, bns = 6, ls = 6,
                     1, 2, pa = .35,  t = 2, b = 2, l = 2, r = 2,
                     c("Old", "Mid_Age", "Young"))
   dev.off()
   png(file = here("Plots", "label-proto-ex3.png"), width=w, height=h)
      make.fuzz.plot(row.lab.dat, col.lab.dat, sup.lab.dat, bns = 6, ls = 6,
                     1, 2, pa = .35,  t = 2, b = 2, l = 2, r = 2,
                     c("Upper_Class", "Lower_Class", "Working_Class", "Middle_Class"))
   dev.off()
   png(file = here("Plots", "label-proto-ex4.png"), width=w, height=h)
      make.fuzz.plot(row.lab.dat, col.lab.dat, sup.lab.dat, bns = 6, ls = 6,
                     1, 2, pa = .35,  t = 2, b = 2, l = 2, r = 2,
                     c("College", "No_College"))
   dev.off()

   png(file = here("Plots", "label-proto-ex5.png"), width=w, height=h)
      make.fuzz.plot(row.lab.dat, col.lab.dat, sup.lab.dat, bns = 6, ls = 6,
                     1, 2, pa = .35,  t = 2, b = 2, l = 2, r = 2,
                     c("Male", "Female"))
   dev.off()
```

>>>>

```{r setup, include=FALSE}
   library(backbone)
   library(conflicted)
   library(dplyr)
   library(FactoMineR)
   library(factoextra)
   library(ggplot2)
   library(ggrepel)
   library(corrplot)
   library(ggridges)
   library(haven)
   library(here)
   library(paletteer)
   library(stringr)
   cols <- paletteer_dynamic("cartography::multi.pal", 20)
   conflicts_prefer(dplyr::filter)
   conflicts_prefer(dplyr::select)
   source(here("Functions", "make.fuzz.plot.R"))
   source(here("Functions", "make.demog.plot.R"))
   source(here("Functions", "demog.dat.R"))
   source(here("Functions", "sdsm.back.R"))
```

```{r import data and select genre stereotype variables}
   #importing data
   dat <- read_dta("C:/Users/Omar Lizardo/Google Drive/MISC DATA SOURCES/SSI-2012/SSI2012.dta")
   #replacing ids with character names
   dat <- bind_cols(tibble(id_char = paste("id", sapply(dat[, 1], as.character), sep = "_")), dat) %>% 
   select(id_char, classicalfemale:hvymtluc, -ends_with("none")) %>%  #selecting relevant columns
   na.omit() #removing missing values
      r <- c("Classical", "Opera", "Jazz", "Broadway", "Easy", "Big_Band", "Classic_Rock", "Country", "Bluegrass", "Folk", "Gospel", "Latin", "Rap", "Blues", "Reggae", "Pop", "Contemp_Rock", "Alt_Rock", "Dance", "Metal")
       c <- c("Female", "Male", "White", "Black", "Hispanic", "Asian", "College", "No_College", "Young", "Mid_Age", "Old", "Lower_Class", "Working_Class", "Middle_Class", "Upper_Class")
```

```{r Summary Table of Averaged Perceptions}
   tab <- colSums(dat[,2:301])/nrow(dat)
   tab <- matrix(tab, nrow = 20,  byrow = TRUE)
   colnames(tab) <- c
   rownames(tab) <- r
   png(file = here("Plots", "main.png"), width=1000, height=700)
   corrplot(tab, is.corr = FALSE, method = "color", 
      tl.col = "black", tl.cex = 1, tl.srt = 45,
      addCoef.col = "black", cl.pos = "n", col.lim = c(0, max(tab)),
      number.cex = 1, number.digits = 2, addgrid.col = "darkgray",
      col=colorRampPalette(c("lightblue","orange","red"))(200))
   dev.off()
```

```{r Creating individual matrices}
   make.mat <- function(x) {
      x <- matrix(as.numeric(x), 20, 15, byrow = TRUE)
      return(x)
      }
   bb.sdsm <- function(x) {
     tryCatch({
         w <- round(sdsm(x, alpha = NULL, missing.as.zero = TRUE)$Pupper, 2)
         return(w)
            }, error = function(e) {return(NULL)})
         }
   den <- rowSums(dat[,2:301])/300
   bad.den <- which(den > .9 | den < 0.1)
   dat2 <- dat[-c(bad.den), 2:301]
   ind.list <- split(dat2, seq(nrow(dat2))) # splitting each row of data frame into list objects
   ind.mat <- lapply(ind.list, make.mat) # creating individual two mode network of genre by labels
   ind.mat <- lapply(ind.mat, function(x) {rownames(x) <- r; x}) # naming individual rows
   ind.mat <- lapply(ind.mat, function(x) {colnames(x) <- c; x}) # naming individual columns
   names(ind.mat) <- dat$id_char[-c(bad.den)]
   ind.mat.t <- lapply(ind.mat, function(x) {t(x)})
   ind.row <- lapply(ind.mat, function(x) {x %*% t(x)}) # individual genre projection
   ind.col <- lapply(ind.mat, function(x) {t(x) %*% x}) # individual label projection
```
   # row.bb <- lapply(ind.mat, bb.sdsm)
   # col.bb <- lapply(ind.mat.t, bb.sdsm)
   # null.rows <- which(as.numeric(unlist(lapply(row.bb, is.null))) == 1)
   # null.cols <- which(as.numeric(unlist(lapply(col.bb, is.null))) == 1)
   # row.bb <- row.bb[-null.rows]
   # col.bb <- col.bb[-null.cols]
   # row.bb <- lapply(row.bb, function(x) {ifelse(is.nan(x), 1, x)})
   # col.bb <- lapply(col.bb, function(x) {ifelse(is.nan(x), 1, x)})
   # row.bb <- lapply(row.bb, function(x) {diag(x) <- 1; x})
   # col.bb <- lapply(col.bb, function(x) {diag(x) <- 1; x})
   # save(row.bb, col.bb, file = here("bb.RData"))
   load(here("bb.RData"))
   p.val <- 0.45
   row.bb <- lapply(row.bb, function(x) {(x < p.val) * 1})
   col.bb <- lapply(col.bb, function(x) {(x < p.val) * 1})
   row.stack <- matrix(unlist(row.bb), ncol = 20, byrow = TRUE) # genre stacked matrix
   col.stack <- matrix(unlist(col.bb), ncol = 15, byrow = TRUE) # label stacked matrix
   colnames(row.stack) <- r # naming genre stacked matrix columns
   colnames(col.stack) <- c # naming label stacked matrix columns
   m <- nrow(row.stack)/20
   rownames(row.stack) <- rep(r, m) # naming genre stacked matrix rows
   rownames(col.stack) <- rep(c, m) # naming label stacked matrix rows
   zero.rows.r <- which(rowSums(row.stack) == 0) # which genre stacked matrix rows sum to zero?
   zero.rows.c <- which(rowSums(col.stack) == 0) # which label stacked matrix rows sum to zero?
```

```{r Stacked CA}
   #genres
   row.stack.d <- data.frame(row.stack)
   row.stack.d$genre <- rep(r, m)
   row.stack.d$id <- rep(names(row.bb), each = 20)
   CA.r <- CA(row.stack.d, graph = FALSE, row.sup = zero.rows.r, quali.sup = c(21, 22))
   #labels
   col.stack.d <- data.frame(col.stack)
   col.stack.d$label <- rep(c, m)
   col.stack.d$id <- rep(names(col.bb), each = 15)
   CA.c <- CA(col.stack.d, graph = FALSE, row.sup = zero.rows.c, quali.sup = c(16, 17))
```

@@

```{r Genre Cloud Plots}
   h <- 625
   w <- 750
   png(file = here("Plots", "genre-proto-ex1.png"), width=w, height=h)
      make.fuzz.plot(row.gen.dat, col.gen.dat, sup.gen.dat, 
                     seed = 125, ls = 5, 
                     1, 2, pa = .35,  t = 3, b = 3, l = 2, r = 3, bns = 6,
                     c("Rap", "Folk", "Classical"))
   dev.off()
   png(file = here("Plots", "genre-proto-ex2.png"), width=w, height=h)
      make.fuzz.plot(row.gen.dat, col.gen.dat, sup.gen.dat,
                     seed = 125, ls = 5,
                     1, 2, pa = .35,  t = 2, b = 2, l = 2, r = 3, bns = 6,
                     c("Alt_Rock", "Gospel", "Opera"))
   dev.off()
   png(file = here("Plots", "genre-proto-ex3.png"), width=w, height=h)
      make.fuzz.plot(row.gen.dat, col.gen.dat, sup.gen.dat,
                     seed = 125, ls = 5,
                     1, 2, pa = .35,  t = 2, b = 2, l = 2, r = 3, bns = 6,
                     c("Metal", "Bluegrass", "Broadway"))
   dev.off()
   png(file = here("Plots", "genre-proto-ex4.png"), width=w, height=h)
      make.fuzz.plot(row.gen.dat, col.gen.dat, sup.gen.dat,
                     seed = 125, ls = 5,
                     1, 2, pa = .35,  t = 2, b = 2, l = 2, r = 3.25, bns = 6,
                     c("Jazz", "Reggae", "Blues", "Rap", "Classical"))
   dev.off()
   png(file = here("Plots", "genre-proto-ex5.png"), width=w, height=h)
      make.fuzz.plot(row.gen.dat, col.gen.dat, sup.gen.dat,
                     seed = 125, ls = 5,
                     1, 2, pa = .35,  t = 2, b = 2, l = 2, r = 3, bns = 6,
                     c("Country", "Bluegrass", "Classic_Rock", "Folk"))
   dev.off()
   png(file = here("Plots", "genre-proto-ex6.png"), width=w, height=h)
      make.fuzz.plot(row.gen.dat, col.gen.dat, sup.gen.dat,
                     seed = 125, ls = 5,
                     1, 2, pa = .35,  t = 2, b = 2, l = 2, r = 3, bns = 6,
                     c("Alt_Rock", "Contemp_Rock", "Classic_Rock", "Metal", "Broadway"))
   dev.off()
   png(file = here("Plots", "genre-proto-ex7.png"), width=w, height=h)
      make.fuzz.plot(row.gen.dat, col.gen.dat, sup.gen.dat,
                     seed = 125, ls = 5,
                     1, 2, pa = .35,  t = 2, b = 2, l = 2, r = 3, bns = 6,
                     c("Broadway", "Classical", "Opera", "Easy", "Jazz"))
   dev.off()
```

```{r Label Cloud Plots}
   h <- 625
   w <- 750
   png(file = here("Plots", "label-proto-ex1.png"), width=w, height=h)
      make.fuzz.plot(row.lab.dat, col.lab.dat, sup.lab.dat, bns = 8, ls = 6,
                     1, 2, pa = .35,  t = 2.5, b = 2, l = 1.5, r = 2.15,
                     c("White", "Asian", "Black", "Hispanic"))
   dev.off()
   png(file = here("Plots", "label-proto-ex2.png"), width=w, height=h)
      make.fuzz.plot(row.lab.dat, col.lab.dat, sup.lab.dat, bns = 6, ls = 6,
                     1, 2, pa = .35,  t = 2.2, b = 2.2, l = 1.5, r = 2.5,
                     c("Old", "Mid_Age", "Young"))
   dev.off()
   png(file = here("Plots", "label-proto-ex3.png"), width=w, height=h)
      make.fuzz.plot(row.lab.dat, col.lab.dat, sup.lab.dat, bns = 6, ls = 6,
                     1, 2, pa = .35,  t = 2, b = 2, l = 2, r = 2,
                     c("Upper_Class", "Lower_Class", "Working_Class", "Middle_Class"))
   dev.off()
   png(file = here("Plots", "label-proto-ex4.png"), width=w, height=h)
      make.fuzz.plot(row.lab.dat, col.lab.dat, sup.lab.dat, bns = 6, ls = 6,
                     1, 2, pa = .35,  t = 2, b = 2, l = 2, r = 2,
                     c("College", "No_College"))
   dev.off()

   png(file = here("Plots", "label-proto-ex5.png"), width=w, height=h)
      make.fuzz.plot(row.lab.dat, col.lab.dat, sup.lab.dat, bns = 6, ls = 6,
                     1, 2, pa = .35,  t = 3, b = 2, l = 2, r = 2,
                     c("Male", "Female"))
   dev.off()
```

### 2025-06-21 Slides/heidelberg-slides.qmd
reorg

---
title: "Musical Genres and Socio-Demographic Labels as Fuzzy Categories"
subtitle: ""
author: "Omar Lizardo"
institute: "*University of California, Los Angeles*"
date: "`r Sys.Date()`"
format:
  revealjs:
    incremental: false 
    theme: simple 
    smaller: true
    slide-number: true
    slide-level: 2
    progress: true
    footer: "https://github.com/olizardo/cultural-consensus-musical-genres"
---

@@

- Notion of Genre Fundamental to the Sociology of Taste
- Perennially Contested Category
- Basic Problem:
  - Our intuitions imply **fuzziness** and **distributional** thinking
  - But our methods mostly rely on **crisp** classifications and **central tendencies**

@@

- Notion of **Genre** Fundamental to the Sociology of Taste
- Perennially Contested Category
- Recent work on **Measuring Culture** has begun to make headway on the issue
  - Relational approaches applying network methods
    - Focus on fuzziness and overlapping categories
  - **Geometric Data Analysis**
    - Focus on distributions of judgments in a relational space

@@

- Recent conceptual and methodological work uses relational imagery to deconstruct categories as clusters of associations between elements
  - **Relational Class Analysis**
  - **Correlational Class Analysis**
  - Measuring **Schemas** from Survey and Discourse Data
- Folk categories are revealed by measuring "what goes with what" at the level of personal culture
- Focus on *heterogeneity* across clusters of individuals and items

@@

## The sociology of Taste
-  Focused on the dual linkage between two systems of categories:
  - Genre categories developed in fields of cultural production (scenes, industries)
  - Social categories endowed with ritual potency and social validity constitutive of status orders 
- Basic intuition:
  - Genres are defined by the categories of people who choose them
  - Categories of persons are defined by the genres they choose
- Duality!

@@

## The sociology of Taste
- We can apply lessons learned from the **relational**, **geometric**, and **schematic** turns in measuring culture to the problem of genre in the sociology of taste
- Investigate folk schemas regarding the (explicit) associations between genre categories and categories of people
- Focus on heterogeneity and the distribution of judgments across the social space

## Data
- Data on cultural tastes among a weighted representative sample of Americans collected in Summer of 2012
  - $N = 2,250$
- Like GSS 1993 survey included items assessing respondents’ likes and dislikes for 20 categories of musical style**
  -  Inclusion of a **perceptual** module
  - Which of these characteristics describe the typical fans of [genre category] music? Choose all that apply.

@@

## Approach
- We can treat the association data provided by respondents as **cognitive two-mode data**
  - In analogy with Krackhardt-style data for one-mode networks
  - Persons provide *perceived* associations between **musical genres** and **social labels**
- Use relational techniques to extract the "backbone" of the perceived associations
  - Dual Projection**
  - Backbone extraction***

@@

## Approach
- We can treat the association data provided by respondents as **cognitive two-mode data**
  - In analogy with Krackhardt-style data for one-mode networks
  - Persons provide *perceived* associations between **musical genres** and **social labels**
- Model the backbone of the person-specific projected networks jointly using: 
  - Stacked **Correspondence Analysis****
  - Geometric Data Analysis of the various **clouds of individuals*****

@@

## Stacked CA Analyses
- Three sets of **scores**:
  - row-scores are **person-specific** judgments of the relative similarity of genres and labels
  - column-scores represent the **aggregate judgment** of the relative similarity of genres (with respect to labels) and labels (with respect to genres)
  - Supplementary scores for genres and labels represent the **centroid** of the cloud of person-specific judgments

## Stacked CA Analyses
- Three sets of **scores**:
  - row-scores are **person-specific** judgments of the relative similarity of genres and labels
  - column-scores represent the **aggregate judgment** of the relative similarity of genres (with respect to labels) and labels (with respect to genres)
  - Supplementary scores for genres and labels represent the **centroid** of the cloud of person-specific judgments
- Multiple **clouds of individuals**
  - Projected into genre space (one cloud per genre)
  - Projected into label space (one cloud per label)
    - Individual specific clouds (across all genres and labels)
    - Sub-clouds in each space based on individual characteristics

- Each individual provides a "personal" two mode network of genres by social labels with corresponding affiliation matrix $\mathbf{A}(p)$
  - $a^{(p)}_{gl} = 1$ if individual $p$ associates genre $g$ with label $l$
- The personal genre projection is given by:
  - $\mathbf{G}(p) = \mathbf{A}(p)\mathbf{A}(p)^T$
  - $g^{(p)}_{ij}$ records the $p^{th}$ individual's perceived similarity between genres $i$ and $j$ based on their shared labels
  - $g^{(p)}_{ii}$ records the number of times individual $p$ associates genre $i$ with a label

@@

>>>>

---
title: "Musical Genres and Socio-Demographic Labels as Fuzzy Categories"
subtitle: ""
author: "Omar Lizardo"
institute: "*University of California, Los Angeles*"
date: "`r Sys.Date()`"
format:
  revealjs:
    embed-resources: true
    incremental: false 
    theme: simple 
    smaller: true
    slide-number: true
    slide-level: 2
    progress: true
    footer: "https://github.com/olizardo/cultural-consensus-musical-genres"
---

@@

- Notion of **genre** fundamental to the Sociology of Taste
- Key organizing principle
- Perennially **contested** Category
- Basic Problem:
  - Our intuitions imply **fuzziness** and **distributional** thinking
  - But our methods mostly rely on **crisp** classifications and **central tendencies**
  - Calls to either reject or radically revise the notion

@@

- Notion of **genre** fundamental to the Sociology of Taste
- Key organizing principle
- Perennially **contested** Category
- Recent work on **Measuring Culture** has begun to make headway on the issue
  - **Relational** approaches applying network methods
    - Focus on fuzziness and overlapping categories
  - **Geometric Data Analysis**
    - Focus on distributions of judgments in a relational space

@@

- Recent conceptual and methodological work uses relational imagery to deconstruct categories as clusters of associations between elements
  - **Relational Class Analysis**
  - **Correlational Class Analysis**
  - Measuring **Schemas** from Survey and Discourse Data
- Folk categories are revealed by measuring **what goes with what** at the level of personal culture
- Focus on **heterogeneity** across clusters of individuals and items

@@

## The sociology of Taste
-  Focused on the objective **dual linkage** between two systems of categories:
  - **Genre categories** developed in fields of cultural production (scenes, industries)
  - **Social categories** endowed with ritual potency and social validity constitutive of status orders 
- Basic intuition:
  - Genres are defined by the categories of people who choose them
  - Categories of persons are defined by the genres they choose
- Yet examining **folk construals** of these dual linkages not a common focus

@@

## Questions
- We can apply lessons learned from the **relational**, **geometric**, and **schematic** turns in measuring culture to the problem of genre in the sociology of taste
  - From **crisp** definitions to **fuzzy** categories
  - Focus on **heterogeneity** and the distribution of judgments across the social space
- We can also move beyond the exclusive focus on objective linkages between genre categories (and cultural objects) and social demographic position 
  - Investigate folk schemas regarding the **subjective folk construals** of the link between genre categories and categories of people
  - Need data containing perceived linkages between genre and social **categories**

## Data
- Data on cultural tastes among a (weighted) representative sample of Americans collected in Summer of 2012
  - $N = 2,250$
- Like GSS 1993 survey included items assessing respondents’ likes and dislikes for 20 categories of musical style**
  -  Inclusion of a **perceptual** module
  - Which of these characteristics describe the typical fans of [genre category] music? Choose all that apply.

@@

## Approach
- We can treat the association data provided by respondents as **cognitive two-mode data**
  - In analogy with Krackhardt-style data for one-mode networks
  - Persons provide *perceived* associations between **musical genres** and **social labels**
- Use relational techniques to extract the **backbone** of the perceived associations isolating the most important ones:
  - Dual Projection**
  - Binarizing and via Backbone extraction***

@@

## Approach
- We can treat the association data provided by respondents as **cognitive two-mode data**
  - In analogy with Krackhardt-style data for one-mode networks
  - Persons provide *perceived* associations between **musical genres** and **social labels**
- Model the backbone of the person-specific projected networks jointly using: 
  - Stacked **Correspondence Analysis****
  - Geometric Data Analysis of the various **clouds of individuals*****
  - Regression, clustering and other methods

@@

## Stacked CA Analyses
- Three sets of **scores**:
  - **Row scores** are **person-specific** judgments of the relative similarity of genres and labels
  - **Column scores** represent the **aggregate judgment** of the relative similarity of genres (with respect to labels) and labels (with respect to genres)
  - **Supplementary scores** for genres and labels represent the **centroid** of the cloud of person-specific judgments
- We can examine heterogeneity in folk understandings of genre and social categories by looking at how the multiple **cloud of individuals** distribute themselves along the **principal axes**

## Stacked CA Analyses
- Three sets of **scores**:
  - **Row scores** are **person-specific** judgments of the relative similarity of genres and labels
  - **Column scores** represent the **aggregate judgment** of the relative similarity of genres (with respect to labels) and labels (with respect to genres)
  - **Supplementary scores** for genres and labels represent the **centroid** of the cloud of person-specific judgments
- Multiple **clouds of individuals**:
  - Projected into **genre space** (one cloud per genre)
  - Projected into **social label space **(one cloud per label)
    - Individual specific clouds (across all genres and labels)
    - Sub-clouds in each space based on individual characteristics

## Mondo Breiger Approach**
- Each individual provides a "personal" two mode network of genres by social labels with corresponding affiliation matrix $\mathbf{A}(p)$
  - $a^{(p)}_{gl} = 1$ if individual $p$ associates genre $g$ with label $l$
- The personal genre projection is given by:
  - $\mathbf{G}(p) = \mathbf{A}(p)\mathbf{A}(p)^T$
  - $g^{(p)}_{ij}$ records the $p^{th}$ individual's perceived similarity between genres $i$ and $j$ based on their shared labels
  - $g^{(p)}_{ii}$ records the number of times individual $p$ associates genre $i$ with a label

@@

## Group-Specific Genre Clouds
:::: {.columns}
::: {.column width="55%"}
![](Plots/demog-college-educ.png)
:::
::: {.column width="45%"}
![Variance of Age Clouds](Plots/demog-college-educ-var.png)
:::
::::

## Group-Specific Genre Clouds
:::: {.columns}
::: {.column width="55%"}
![](Plots/demog-nocollege-educ.png)
:::
::: {.column width="45%"}
![Variance of Age Clouds](Plots/demog-nocollege-educ-var.png)
:::
::::

## Discussion and Conclusion
- Exploiting the sociological duality of **genres** and **social labels** to extract **associational schemas** for both

- Combining **relational** and **geometric** methods to examine central tendencies (*centroids*) and heterogeneity (*variance*) simultaneously
- Move across multiple **levels**
  - Individual (typical/atypical)
  - Category
  - Group (Demographics)
  - System
- Examining variation in **meaning consensus** across levels

## Final Slide

- Thanks!

:::aside
olizardo@soc.ucla.edu
:::
